---
title: "Inflation by State"
author: "Tara Hoffman"
date: "2024-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://www.bls.gov/regions/subjects/consumer-price-indexes.htm#AZ


New York: 8,804,190
Los Angeles: 3,898,747
Chicago: 2,746,388
Houston: 2,304,580
Phoenix: 1,608,139
Philadelphia: 1,603,797
San Antonio: 1,434,625
San Diego: 1,386,932
Dallas: 1,304,379
San Jose: 1,013,240



Largest Metropolitan Areas: 
https://www.thoughtco.com/largest-metropolitan-areas-1435135

1.New York
2.Los Angelos
3.Chicago  
4.Dallas + Fort Worth- NO Data
5. Houston - No Data
6. DC 
7. Miami + Ft.Lauderdale
8. Philadelphia 
9. Atlanta
10. Pheonix 



New York:
```{r}


#getting 12 month average in each year
rvest:: read_html_live("https://www.bls.gov/regions/northeast/news-release/consumerpriceindex_newyork.htm") |> rvest::html_element("#CPITableANewYork\\.xlsx > tbody") |> rvest::html_table() -> newyork

colnames(newyork)<- c("Month", "2020 3mo", "2020 12mo", "2021 3mo", "2021 12mo", "2022 3mo", "2022 12mo", "2023 3mo", "2023 12mo", "2024 3mo", "2024 12mo")

newyork$city<- "New York City"

newyork<-newyork[,-c(2,4,6,8,10)]

colnames(newyork)<- c("Month","2020", "2021", "2022","2023","2024", "City")

newyork <- newyork[, c("City", "Month", "2020", "2021","2022", "2023", "2024")]

newyork<- tidyr::pivot_longer(
  
  data= newyork, 
  cols = "2020":"2024", 
  names_to = "Year", 
  values_to = "12 mo Avg Inflation"
)




```



Los Angeles:
```{r}

#Los Angeles


#getting 12 month average in each year
rvest:: read_html_live("https://www.bls.gov/regions/west/news-release/consumerpriceindex_losangeles.htm") |> rvest::html_elements("#CPITableALosAngeles\\.xlsx > tbody")|> rvest:: html_table() -> losangeles

losangeles<- losangeles[[1]]

colnames(losangeles)<- c("Month", "2020 3mo", "2020 12mo", "2021 3mo", "2021 12mo", "2022 3mo", "2022 12mo", "2023 3mo", "2023 12mo", "2024 3mo", "2024 12mo")

losangeles$city<- "Los Angeles"

losangeles<-losangeles[,-c(2,4,6,8,10)]

colnames(losangeles)<- c("Month","2020", "2021", "2022","2023","2024", "City")

losangeles<- losangeles[, c("City", "Month", "2020", "2021","2022", "2023", "2024")]

losangeles<- tidyr::pivot_longer(
  
  data= losangeles, 
  cols = "2020":"2024", 
  names_to = "Year", 
  values_to = "12 mo Avg Inflation"
)


```



Chicago: 

- has revised data notated with (r)
```{r}
#getting 12 month average in each year
rvest:: read_html_live("https://www.bls.gov/regions/midwest/news-release/consumerpriceindex_chicago.htm") |> rvest::html_elements("#CPITableAChicago\\.xlsx > tbody")|> rvest:: html_table() -> chicago

chicago <- chicago[[1]]

colnames(chicago)<- c("Month", "2020 3mo", "2020 12mo", "2021 3mo", "2021 12mo", "2022 3mo", "2022 12mo", "2023 3mo", "2023 12mo", "2024 3mo", "2024 12mo")

chicago$city<- "Chicago"

chicago<-chicago[,-c(2,4,6,8,10)]

colnames(chicago)<- c("Month","2020", "2021", "2022","2023","2024", "City")

chicago<- chicago[, c("City", "Month", "2020", "2021","2022", "2023", "2024")]

#replacing revised data(r) with numeric


chicago [6,5]<- "8.9"
chicago [7,5]<- "8.4"
chicago [8,5]<- "8.3" 
chicago [9,5]<- "8.2"
chicago [10,5]<-"7.7"

chicago$`2022` <- as.numeric(chicago$`2022`)

chicago<- tidyr::pivot_longer(
  
  data= chicago, 
  cols = "2020":"2024", 
  names_to = "Year", 
  values_to = "12 mo Avg Inflation"
)

#delete rows 61-65
chicago<- chicago[-c(61:65),]


```


DC - Data only collected every other month 
```{r}



#getting 12 month average in each year
rvest:: read_html_live("https://www.bls.gov/regions/mid-atlantic/news-release/consumerpriceindex_washingtondc.htm") |> rvest::html_elements("#CPITableAWashington\\.xlsx > tbody")|> rvest:: html_table() -> dc

dc<- dc[[1]]

colnames(dc)<- c("Month", "2020 3mo", "2020 12mo", "2021 3mo", "2021 12mo", "2022 3mo", "2022 12mo", "2023 3mo", "2023 12mo", "2024 3mo", "2024 12mo")

dc$city<- "DC"

dc<-dc[,-c(2,4,6,8,10)]

colnames(dc)<- c("Month","2020", "2021", "2022","2023","2024", "City")

dc<- dc[, c("City", "Month", "2020", "2021","2022", "2023", "2024")]

dc<- tidyr::pivot_longer(
  
  data= dc, 
  cols = "2020":"2024", 
  names_to = "Year", 
  values_to = "12 mo Avg Inflation"
)


```


Miami Ft. Lauderdale:
```{r}

#getting 12 month average in each year
rvest:: read_html_live("https://www.bls.gov/regions/southeast/news-release/consumerpriceindex_miami.htm") |> rvest::html_elements("#CPITableAMiami\\.xlsx > tbody")|> rvest:: html_table() -> miami

miami<- miami[[1]]

colnames(miami)<- c("Month", "2020 3mo", "2020 12mo", "2021 3mo", "2021 12mo", "2022 3mo", "2022 12mo", "2023 3mo", "2023 12mo", "2024 3mo", "2024 12mo")

miami$city<- "Miami"

miami<-miami[,-c(2,4,6,8,10)]

colnames(miami)<- c("Month","2020", "2021", "2022","2023","2024", "City")

miami<- miami[, c("City", "Month", "2020", "2021","2022", "2023", "2024")]

miami<- tidyr::pivot_longer(
  
  data= miami, 
  cols = "2020":"2024", 
  names_to = "Year", 
  values_to = "12 mo Avg Inflation"
)

```


Philadelphia: 

```{r}

#getting 12 month average in each year
rvest:: read_html_live("https://www.bls.gov/regions/mid-atlantic/news-release/consumerpriceindex_philadelphia.htm") |> rvest::html_elements("#CPITableAPhiladelphia\\.xlsx > tbody")|> rvest:: html_table() -> philadelphia

philadelphia<- philadelphia[[1]]

colnames(philadelphia)<- c("Month", "2020 3mo", "2020 12mo", "2021 3mo", "2021 12mo", "2022 3mo", "2022 12mo", "2023 3mo", "2023 12mo", "2024 3mo", "2024 12mo")

philadelphia$city<- "Philadelphia"

philadelphia<-philadelphia[,-c(2,4,6,8,10)]

colnames(philadelphia)<- c("Month","2020", "2021", "2022","2023","2024", "City")

philadelphia<- philadelphia[, c("City", "Month", "2020", "2021","2022", "2023", "2024")]

philadelphia<- tidyr::pivot_longer(
  
  data= philadelphia, 
  cols = "2020":"2024", 
  names_to = "Year", 
  values_to = "12 mo Avg Inflation"
)

```



Atlanta: 

```{r}
#getting 12 month average in each year
rvest:: read_html_live("https://www.bls.gov/regions/southeast/news-release/consumerpriceindex_atlanta.htm") |> rvest::html_elements("#CPITableAAtlanta\\.xlsx > tbody")|> rvest:: html_table() -> atlanta

atlanta<- atlanta[[1]]

colnames(atlanta)<- c("Month", "2020 3mo", "2020 12mo", "2021 3mo", "2021 12mo", "2022 3mo", "2022 12mo", "2023 3mo", "2023 12mo", "2024 3mo", "2024 12mo")

atlanta$city<- "Atlanta"

atlanta<-atlanta[,-c(2,4,6,8,10)]

colnames(atlanta)<- c("Month","2020", "2021", "2022","2023","2024", "City")

atlanta<- atlanta[, c("City", "Month", "2020", "2021","2022", "2023", "2024")]

atlanta<- tidyr::pivot_longer(
  
  data= atlanta, 
  cols = "2020":"2024", 
  names_to = "Year", 
  values_to = "12 mo Avg Inflation"
)

```


Anchorage: 

```{r}
#getting 12 month average in each year
rvest:: read_html_live("https://www.bls.gov/regions/west/news-release/consumerpriceindex_anchorage.htm") |> rvest::html_elements("#tableA\\.xlsx > tbody")|> rvest:: html_table() -> anchorage

anchorage<- anchorage[[1]]

colnames(anchorage)<- c("Month", "2020 3mo", "2020 12mo", "2021 3mo", "2021 12mo", "2022 3mo", "2022 12mo", "2023 3mo", "2023 12mo", "2024 3mo", "2024 12mo")

anchorage$city<- "Anchorage"

anchorage<-anchorage[,-c(2,4,6,8,10)]

colnames(anchorage)<- c("Month","2020", "2021", "2022","2023","2024", "City")

anchorage<- anchorage[, c("City", "Month", "2020", "2021","2022", "2023", "2024")]

anchorage<- tidyr::pivot_longer(
  
  data= anchorage, 
  cols = "2020":"2024", 
  names_to = "Year", 
  values_to = "12 mo Avg Inflation"
)
```






By Region: 

Mountain Plains: Colorado, Kansas, Missouri, Montana, Utah, Wyoming

South

West: Alaska, Arizona, California, Guam, Hawaii, Idaho, Nevada, Oregon, Washington
NorthEast
Midwest

South east: Alabama, Florida, Georgia, Kentucky, Mississippi, North Carolina, South Carolina, Tennessee

Mid Atlantic: Delaware, District of Columbia, Maryland, New Jersey, Pennsylvania, Virginia, West Virginia

Midwest


Midwest:

```{r}
#getting 12 month average in each year
rvest:: read_html_live("https://www.bls.gov/regions/mountain-plains/news-release/consumerpriceindex_midwest.htm") |> rvest::html_elements("#tablea > tbody")|> rvest:: html_table() -> midwest

midwest<- midwest[[1]]

colnames(midwest)<- c("Month", "2020 3mo", "2020 12mo", "2021 3mo", "2021 12mo", "2022 3mo", "2022 12mo", "2023 3mo", "2023 12mo", "2024 3mo", "2024 12mo")

midwest$Region<- "Midwest"

midwest<-midwest[,-c(2,4,6,8,10)]

colnames(midwest)<- c("Month","2020", "2021", "2022","2023","2024", "Region")

midwest<- midwest[, c("Region", "Month", "2020", "2021","2022", "2023", "2024")]

midwest<- tidyr::pivot_longer(
  
  data= midwest, 
  cols = "2020":"2024", 
  names_to = "Year", 
  values_to = "12 mo Avg Inflation"
)


```

https://www.bls.gov/regions/west/news-release/consumerpriceindex_west.htm

Western: 





